The LSST DESC 3x2pt Tomography Optimization Challenge

### 2021-02-17 paper/paper.tex
fixes
new text and placeholder plots

In this paper we discuss and evaluate methods for assigning objects to tomographic bins, and 
specifically the impact
of using a limited set of color bands to make those assignments, motivated by requirements of using the
metacalibration method for galaxy shape measurement to limit biases.  Tomographic methods have been 
explored before in \citet{jain} and \citet{kitching2019}.  In this paper we describe the results of a 
challenge using simulated Rubin-like data with a range of methods submitted by different teams.  We explore how many tomographic bins we can effectively generate using only a subset of bins, and compare methods submitted to the challenge.

@@

@@

@@

\begin{figure}[htbp]
% \centering
\includegraphics[width=1\linewidth]{results/funbins_nz.pdf}
\caption{$n(z)$ values from funbins}
\label{fig:funbin_nz}
\end{figure}

@@

\begin{enumrate}
\item Generally good relationship, especially at high end
\item Domination of metric by clustering
\item Bimodality in SNR
\end{enumerate}

@@

\subsection{Trained edges vs fixed edges}

@@

\begin{figure*}
\includegraphics[width=0.9\columnwidth]{results/edge_comparison.pdf}
\caption{A comparison of results from methods which used fixed target edges (dashed blue)
with those which optimized target bins (solid orange).  Methods were trained on the 3x2pt metric
shown in the upper panel; the lower performance for these classifications 
for the lensing-only metric shown in the lower panel illustrates the need for science case-specific
binning choices. In each case, classification is done using griz bands on the CosmoDC2 data set.}
\label{fig:edges}
\end{figure*}

@@

\subsection{Future directions}

\begin{itemize}
    \item spectroscopic incompleteness / unrepresentativeness
    \item data set volume
    \item comparison of optimal choices in different science cases
\end{itemize}

@@

\section{Mistakes}
The challenge organizers made a number of mistakes when building and running the challenge.
These do not invalidate our process or results, but we describe some of them here in the hope of
being useful for future challenge designers.

>>>>

In this paper we discuss and evaluate methods for assigning objects to tomographic bins, and 
specifically the impact
of using a limited set of color bands to make those assignments, motivated by requirements of using the
metacalibration method for galaxy shape measurement to limit biases.  Tomographic methods have been 
explored before in \citet{jain} and \citet{kitching2019} \textbf{more}.  In this paper we describe the results of a 
challenge using simulated Rubin-like data with a range of methods submitted by different teams.  We explore how many tomographic bins we can effectively generate using only a subset of bins, and compare methods submitted to the challenge.

@@

The overall number density of the two data sets is shown in Figure \ref{fig:initial_nz},
and a set of colour-colour diagrams  in Figure \ref{fig:colour_colour}.

\begin{figure}[htbp]
\centering
\includegraphics[width=1\linewidth]{results/initial_data.pdf}
\caption{The underlying number density $n(z)$ for the two catalogs used in the challenge,
after the initial cuts described in the text.}
\label{fig:initial_nz}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=1\linewidth]{results/colour_colour.pdf}
\caption{A colour-colour diagram for the two catalogs used in the challenge, showing the bands
available to participants.}
\label{fig:colour_colour}
\end{figure}

@@

A technical problem, involving conflicting versions of libraries and incompatible hardware,
meant that method could not be run in the final challenge.

@@

\begin{figure}[htbp]
% \centering
\includegraphics[width=1\linewidth]{results/funbins_nz.pdf}
\caption{The tomographic $n(z)$ values generated by the winning method in the challenge, {\sc funbins},
generating nine bins. 
The upper panel shows results for the method applied to riz bands, and the lower on griz. The
increased spread and overlap of the bins in the lower panel decreases the overall constraining
power and signal-to-noise.}
\label{fig:funbin_nz}
\end{figure}

@@

The different metrics are generally well-correlated (this also holds true for the metrics not 
shown here), especially at the highest values of the metrics representing the most successful
methods. This is particularly so when comparing the Buzzard and CosmoDC2 metrics, which is encouraging 
as it implies these methods are broadly stable to reasonable changes in the underlying galaxy
distribution, and hence on the survey properties.  The largest outlier comes from the 
{\sc Neural Network} entry.

The relatively broader spread in the lensing-only metric illustrates how our metric is dominated
by the stronger constraining power in the clustering part of the data set.  This arises because
we do not include a galaxy bias model in our FOM, and so the constraining power of the clustering
is artifically high.
However, since the clustering metric then largely becomes a proxy for how well separated
galaxy bins are, this is not a major limitation of the metrics.

Finally, there is a bimodal structure in the relationship between the FOM and SNR metrics.
This largely traces the split between models which train the target bin edges vs those
which fix them, as described below in Section \ref{sec:train-vs-fix}.

@@

\subsection{Trained edges vs fixed edges}\label{sec:train-vs-fix}

@@

\begin{figure}
\includegraphics[width=0.9\columnwidth]{results/edge_comparison.pdf}
\caption{A comparison of results from methods which used fixed target edges (dashed blue)
with those which optimized target bins (solid orange).  Methods were trained on the 3x2pt metric
shown in the upper panel; the lower performance for these classifications 
for the lensing-only metric shown in the lower panel illustrates the need for science case-specific
binning choices. In each case, classification is done using griz bands on the CosmoDC2 data set.}
\label{fig:edges}
\end{figure}

@@

\section{Conclusions}
The Tomography Challenge has proved a useful mechanism for encouraging the development
of methods for this science question.  We encourage this challenge structure for other
well-specified science cases, and DESC has subsequently run other challenges along the same lines
such is its N5K beyond-limber challege\footnote{\url{https://github.com/LSSTDESC/N5K}}.
We describe in Appendix \ref{sec:mistakes} some lessons we have learned from this challenge
that may be useful for future ones.

The results of the challenge have shown first that three or four-band photometry
is sufficient for specifying tomography even up to a large number of bins, if
the training data is complete enough: degneracies between colour do not make this impossible.
Excluding the g-band from the data reduces the constraining power by a noticeable but
not catastrophic amount.

Finally, we have shown that in general there is a signficant advantage to optimizing
target bins afresh when considering a new science case; this can be a significant boost
to the final constraining power.
Despite this, evenly spaced bins in fiducial co-moving distance have proven
a good general choice in the winning method in the challenge, {\sc funbins}.

The challenge was heavily simplified, with the intention of testing whether
effective bin-assignments with limited photometry is possible even in theory,
and of generating as many potential methods for the task as possible.  With
both these acheived, future directions for simulating the problem are clear.
The most obvious is limited spectroscopic completeness and size.  Realistic
training data sets will be far smaller than the millions of objects we used here,
and the difficulty of measuring spectra for faint galaxies will make the them
incomplete and highly unrepresentative. Machine learning methods will be particularly
affected by the latter.  Future challenges must explore these important limitations
using simulated incompleteness.

@@

\section{Mistakes} \label{sec:mistakes}
The challenge organizers made a number of mistakes when building and running the challenge.
These do not invalidate our process or results, but we describe some of them here in the hope of
being useful for future challenge designers.
